---
title: "Antiboitic dispensing practices"
author: "Abdi Egeh"
date: "`r Sys.Date()`"
output:
 html_document:
    code_download: true 
    code_folding: show
    toc: true 
    toc_float: true 
    theme: readable 
    highlight: zenburn
---

# Loading and libraries
```{r}
library("meta")
library("metafor")
metdata <- read.csv("metadata.csv")
View(metdata)
```

#Transformations
```{r data.transformations}
metdata$study_design <- factor(metdata$design,
                            levels = c(0, 1),
                            labels = c("questionnaire-based", "simulated"))

#region sub groups

metdata$region <- factor(metdata$region,
                            levels = c(0, 1,2),
                            labels = c("West Africa", "East Africa", "Southern Africa"))

metdata$country <- factor(metdata$country, 
                          levels = c(0,1,2,3,4,5,6,7),
                          labels = c("Ghana","Ethiopia","Kenya","Nigeria",
                                     "South Africa","Tanzania", "Uganda", "Zambia"))

# the levels represents how the data was entered in the data set fields. I coded the data as the following 
#0= questionnaire-based data collection method 
#1= simulated client-based
#2= other.

metdata$sample.size <- factor(
  ifelse(metdata$sample <= 86, 0, 1), # this is where the new values are assigned. if the sample size is less than 100, its assigned 0, if not 1.
  levels = c(0, 1),
  labels = c("small", "large"))

```

# Transformation
```{r}

ies.da <- escalc(xi = URTI, ni = sample , data = metdata , measure = "PFT", add = 0) 
#xi=number of events, ni=total
# Individual estimates with double arcsine transformation. This extracts the variance from the two. this is the same as the above 

pes.da <- rma(yi , vi , data = ies.da) # fitting the mixed-effects model here.
#This is the estimated transformed summary proportion.
pes <- predict(pes.da , transf = transf.ipft.hm , 
               targ = list(ni = metdata$sample))
#back fransformation from arcsine to the original.
# displays the pooled estimate and its 95% confidence interval
print(pes.da, digits = 4)
confint(pes.da, digits =4)

#Testing for heteroginiety in the pooled estimate.
```
# Test of moderator analysis
```{r}
mod.design <- rma(yi, vi, mods = ~ factor(study_design), data = ies.da)
mod.design

mod.ssize <- rma(yi, vi, mods = ~ factor(sample.size), data = ies.da)
mod.ssize

```

# Forest plot function (Double arcensine-transformed)
```{r}
pes.summary <- metaprop(event = URTI,
                        n = sample,
                        studlab = Author,
                        data = metdata,
                        sm = "PFT",
                        method.tau = "DL",
                        method.ci = "NAsm")
# a meta-analysis of proportions is conducted here.
```

# Plot visualization (publication-ready)

```{r}
tiff("URTIForest.tiff", width = 12.5, height = 8, units = "in", res = 300)
# saving a plot with a better definition in an image form.
forest(pes.summary,
       sortvar = 1 / pes.summary$seTE,  # Sort by effect size
       comb.random = TRUE,        # Random effects model
       comb.fixed = FALSE,        # Do not show fixed effect
       print.tau2 = TRUE,
       print.I2 = TRUE,
       print.pval.Q = TRUE,
       print.Q = TRUE,
       leftcols = c("studlab", "event", "n", "effect", "ci"),
       leftlabs = c("Study", "Events", "Sample", "Proportion", "95% CI"),
       smlab = "",                # No summary label
       xlab = "% of non-prescriped antibiotics dispensed for URTI",
       backtransf = TRUE,         # Show results in original scale (0 to 1)
       pscale = 100,              # Convert to percentage
       xlim = c(0, 100),          # X-axis range in %
       squaresize = 1,          # Increase box size
       col.square = "navy",
       col.square.lines = "navy",
       col.diamond = "darkred",
       col.diamond.lines = "darkred",
       col.label.left = "black",
       col.label.right = "black",
       fs.study = 11,             # Font size for study labels
       fs.axis = 11,              # Font size for x-axis
       fs.hetstat = 9,            # Font size for heterogeneity stats
       print.subgroup.labels = TRUE,
       spacing = 1.25)             # Adjust line spacing
#dev.off()

```
The results of the forst plot are sorted according to their precision.
# Identifying outliers (Baujta plot)

```{r}
tiff("baujat.dispensing.tiff", width = 12.5, height = 8, units = "in", res = 300)
# Generate the Baujat plot with enhanced visuals
bjplot <- baujat(pes.summary,
                 symbol = 19,                   # Solid circle
                 cex = 1.5,                     # Increase point size
                 col = "navy",                  # Color for points
                 xlim = c(0, 15),
                 ylim = c(0, 0.6),              # Adjust y-axis for space
                 xlab = "Contribution to Overall Heterogeneity",
                 ylab = "Influence on Summary Proportion",
                 main = "Baujat Plot: Influence & Heterogeneity")

# Highlight influential studies
highlight <- bjplot[bjplot$x >= 10 | bjplot$y >= 0.4, ]

# Add labels to those studies
text(highlight$x,
     highlight$y,
     labels = highlight$slab,
     pos = 3,                   # Position above point
     cex = 0.9,                 # Font size
     col = "darkred",
     font = 2)                  # Bold font
dev.off()
```
# Externally standerized residuals (ESR)
```{r}
stud.res <- rstudent(pes.da)
# Compute influence diagnostics

# Sorting ESR by z-values in descending order 
# Get absolute studentized residuals
abs.z <- abs(stud.res$z)

# Order the residuals from largest to smallest
stud.res[order(-abs.z)]
inf <- influence(pes.da) 
print(inf , digits =3)
```
# leave-one-out analysis

# Sub group analysis (meta regresion) defining the column

```{r}

pes.da.subgroup1 <- rma(yi ,vi ,data = ies.da,
                     subset = study_design == "questionnaire-based")
pes.da.subgroup2 <- rma(yi ,vi ,data = ies.da, 
                     subset = study_design == "simulated")

pes.subgroup1 <- predict(pes.da.subgroup1, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.subgroup2 <- predict(pes.da.subgroup2, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))


# dataframe that stores the effect size estimates 
dat.diffvar <- data.frame(estimate = c(pes.da.subgroup1$b,
                                       pes.da.subgroup2$b),
                          stderror = c(pes.da.subgroup1$se ,
                                       pes.da.subgroup2$se), 
                          study_design = c("questionnaire-based", "simulated"), 
                          tau2 = round(c(pes.da.subgroup1$tau2,
                                         pes.da.subgroup2$tau2), 2))
#Fitting a mixed effects regresion to compare the two sub groups
subganal.moderator <- rma(estimate , sei = stderror , 
                          mods = ~ study_design , method = "FE", 
                          data = dat.diffvar)
#Recalculate summary effect size assuming different # heterogeneity components
pes.da.moderator <- rma(estimate , sei = stderror , 
                        method = "FE", data = dat.diffvar)
pes.moderator <- predict(pes.da.moderator , transf = transf.ipft.hm , 
                         targ = list(ni = metdata$sample))

```

#Displaying the results of the subgroup analysis
```{r}
#Subrgoup1 analysis
print(pes.subgroup1); print(pes.da.subgroup1)
#subgroup2 analysis
print(pes.subgroup2); print(pes.da.subgroup2)
# Display subgroup analysis results 
print(subganal.moderator) 
# Display recomputed summary effect size
print(pes.moderator)

```
# Explanation

# forest plot by study design
```{r}

subganal.studydesign <- rma(yi , vi , data = ies.da, 
                            mods = ~ study_design , method = "DL")

pes.summary <- metaprop(URTI, sample , Author,
                        data = metdata , sm = "PLO", method.tau = "DL", 
                        method.ci ="NAsm",
                         byvar = study_design , tau.common=TRUE ,
                        tau.preset = sqrt(subganal.studydesign$tau2))
tiff("dispen_subgroupforestplott.tiff", width = 10, height = 10, units = "in", res = 300)
forest(pes.summary , 
       common = FALSE ,
       overall = TRUE , 
       overall.hetstat = TRUE , 
       resid.hetstat = FALSE , 
       subgroup.hetstat = TRUE , 
       test.subgroup = FALSE , 
       fs.hetstat = 10,
       print.tau2 = TRUE ,
       print.Q = TRUE , 
       print.pval.Q = TRUE , 
       print.I2 = TRUE , 
       rightcols = FALSE , 
       xlim = c(0 ,1),
       leftcols = c("studlab", "effect", "ci"), 
       leftlabs = c("Study", "Proportion", "95% C.I."), 
       text.random.w = "Subgroup proportion", 
       text.random = "Overall proportion", 
       xlab = "Prevalence of CC (%)", 
       pscale = 1, 
       smlab = " ", 
       weight.study = "random", 
       squaresize = 0.5, 
       col.square = "navy", 
       col.diamond = "maroon",
       col.diamond.lines = "maroon", 
       digits = 3,  
       spacing = 1.3)
dev.off()
```



# Sample size moderator 
```{r}
pes.da.small <- rma(yi ,vi ,data = ies.da,
                     subset = sample.size == "small")
pes.da.large <- rma(yi ,vi ,data = ies.da, 
                     subset = sample.size == "large")
pes.small <- predict(pes.da.small, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.large <- predict(pes.da.large, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))

#Data
dat.diffvar <- data.frame(estimate = c(pes.da.small$b,
                                       pes.da.large$b),
                          stderror = c(pes.da.small$se ,
                                       pes.da.large$se), 
                          moderator = c("small", "large"), 
                          tau2 = round(c(pes.da.small$tau2 ,
                                         pes.da.large$tau2), 3))

#Fix effects model

subganal.moderator <- rma(estimate , sei = stderror , 
                          mods = ~ moderator , method = "FE", 
                          data = dat.diffvar)

pes.da.moderator <- rma(estimate , sei = stderror , 
                        method = "FE", data = dat.diffvar)
pes.moderator <- predict(pes.da.moderator , transf = transf.ipft.hm , 
                         targ = list(ni = metdata$sample))



#Subrgoup1 analysis
print(pes.small); print(pes.da.small)
#subgroup2 analysis
print(pes.large); print(pes.da.large)
# Display subgroup analysis results 
print(subganal.moderator) 
# Display recomputed summary effect size
print(pes.moderator)
```

#region sub groups
```{r}

pes.da.subgroup1 <- rma(yi ,vi ,data = ies.da,
                     subset = region == "West Africa")
pes.da.subgroup2 <- rma(yi ,vi ,data = ies.da, 
                     subset = region == "East Africa")
pes.da.subgroup3 <- rma(yi ,vi ,data = ies.da,
                     subset = region == "Southern Africa")
pes.subgroup1 <- predict(pes.da.subgroup1, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.subgroup2 <- predict(pes.da.subgroup2, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.subgroup3 <- predict(pes.da.subgroup3, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))

# dataframe that stores the effect size estimates 
dat.diffvar <- data.frame(estimate = c(pes.da.subgroup1$b,
                                       pes.da.subgroup2$b,
                                       pes.da.subgroup3$b),
                          stderror = c(pes.da.subgroup1$se ,
                                       pes.da.subgroup2$se,
                                       pes.da.subgroup3$se), 
                          region = c("West Africa", "East Africa", "Southern Africa"), 
                          tau2 = round(c(pes.da.subgroup1$tau2 ,
                                         pes.da.subgroup2$tau2,
                                         pes.da.subgroup3$tau2), 3))
#Fitting a mixed effects regresion to compare the two sub groups
subganal.moderator <- rma(estimate , sei = stderror , 
                          mods = ~ region , method = "FE", 
                          data = dat.diffvar)
#Recalculate summary effect size assuming different # heterogeneity components
pes.da.moderator <- rma(estimate , sei = stderror , 
                        method = "FE", data = dat.diffvar)
pes.moderator <- predict(pes.da.moderator , transf = transf.ipft.hm , 
                         targ = list(ni = metdata$sample))
```

# region results display
```{r}
#Subrgoup1 analysis
print(pes.subgroup1); print(pes.da.subgroup1)
#subgroup2 analysis
print(pes.subgroup2); print(pes.da.subgroup2)
print(pes.subgroup3); print(pes.da.subgroup3)
# Display subgroup analysis results 
print(subganal.moderator) 
# Display recomputed summary effect size
print(pes.moderator)
```

#country subgroups

```{r}
pes.da.ghana <- rma(yi ,vi ,data = ies.da,
                     subset = country == "Ghana")
pes.da.ethiopia <- rma(yi ,vi ,data = ies.da, 
                     subset = country == "Ethiopia")
pes.da.kenya <- rma(yi ,vi ,data = ies.da,
                     subset = country == "Kenya")
pes.da.nigeria <- rma(yi ,vi ,data = ies.da,
                     subset = country == "Nigeria")
pes.da.southA <- rma(yi ,vi ,data = ies.da,
                     subset = country == "South Africa")
pes.da.tanzania <- rma(yi ,vi ,data = ies.da,
                     subset = country == "Tanzania")
pes.da.zambia <- rma(yi ,vi ,data = ies.da,
                     subset = country == "Zambia")
pes.ghana <- predict(pes.da.ghana, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.ethiopia <- predict(pes.da.ethiopia, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.kenya <- predict(pes.da.kenya, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.nigeria <- predict(pes.da.nigeria, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.southA <- predict(pes.da.southA, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))
pes.tanzania <- predict(pes.da.tanzania, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))

pes.zambia <- predict(pes.da.zambia, 
                         transf = transf.ipft.hm ,
                         targ = list(ni = metdata$sample))

# Data frame
dat.country <- data.frame(estimate = c(pes.da.ghana$b,
                                       pes.da.ethiopia$b,
                                       pes.da.kenya$b,
                                       pes.da.nigeria$b,
                                       pes.da.southA$b,
                                       pes.da.tanzania$b,
                                       pes.da.zambia$b),
                          stderror = c(pes.da.ghana$se ,
                                       pes.da.ethiopia$se,
                                       pes.da.kenya$se,
                                       pes.da.nigeria$se ,
                                       pes.da.southA$se,
                                       pes.da.tanzania$se,
                                       pes.da.zambia$se), 
                          country = c("Ghana","Ethiopia","Kenya","Nigeria",
                                     "South Africa","Tanzania", "Zambia"), 
                          tau2 = round(c(pes.da.ghana$tau2 ,
                                         pes.da.ethiopia$tau2,
                                         pes.da.kenya$tau2,
                                         pes.da.nigeria$tau2 ,
                                         pes.da.southA$tau2,
                                         pes.da.tanzania$tau2,
                                         pes.da.zambia$tau2), 3))
#Sub group analysis
subganal.moderator.country <- rma(estimate , sei = stderror , 
                          mods = ~ country , method = "FE", 
                          data = dat.country)
#Recalculate summary effect size assuming different # heterogeneity components
pes.da.moderator.country <- rma(estimate , sei = stderror , 
                        method = "FE", data = dat.country)
pes.moderator.country <- predict(pes.da.moderator.country , transf = transf.ipft.hm , 
                         targ = list(ni = metdata$sample))
```

# Print sub country analysis

```{r}
print(subganal.moderator.country) 
# Display recomputed summary effect size
print(pes.moderator.country)
```




# sub region plot 

```{r}
metdata <- metdata[!is.na(metdata$study_design), ]
subganal.regions <- rma(yi , vi , data = ies.da, 
                            mods = ~ region , method = "DL")

pes.summary <- metaprop(advice1 , sample , Author,
                        data = metdata , sm = "PLO", method.tau = "DL", 
                        method.ci ="NAsm",
                        byvar = region , tau.common=TRUE ,
                        tau.preset = sqrt(subganal.regions$tau2))
tiff("consult_regions.tiff", width = 10, height = 10, units = "in", res = 300)
forest(pes.summary , 
       common = FALSE ,
       overall = TRUE , 
       overall.hetstat = TRUE , 
       resid.hetstat = FALSE , 
       subgroup.hetstat = TRUE , 
       test.subgroup = FALSE , 
       fs.hetstat = 11,
       print.tau2 = TRUE ,
       backtransf = TRUE ,
       print.Q = TRUE , 
       print.pval.Q = TRUE , 
       print.I2 = TRUE , 
       rightcols = FALSE , 
       xlim = c(0 ,1),
       leftcols = c("studlab",  "effect",  "ci"), 
       leftlabs = c("Study",  "Proportion",  "95% C.I."), 
       text.random.w = "Subgroup proportion", 
       text.random = "Overall proportion", 
       xlab = "% of people consulted on antibiotic use", 
       pscale = 1, 
       smlab = " ", 
       weight.study = "random", 
       squaresize = 0.5, 
       col.square = "black", 
       col.diamond = "black",
       col.diamond.lines = "black", 
       digits = 2,  
       spacing = 1.3)
dev.off()
```

# Plot- subgroup country 

```{r}
subganal.country <- rma(yi , vi , data = ies.da, 
                            mods = ~ country , method = "DL")

pes.summary.country <- metaprop(advice1 , sample , Author,
                        data = metdata , sm = "PLO", method.tau = "DL", 
                        method.ci ="NAsm",
                        byvar = country , tau.common=TRUE ,
                        tau.preset = sqrt(subganal.country$tau2))
tiff("URTI_country.tiff", width = 10, height = 11.5, units = "in", res = 300)
forest(pes.summary.country , 
       common = FALSE ,
       overall = TRUE , 
       overall.hetstat = FALSE , 
       resid.hetstat = FALSE , 
       subgroup.hetstat = TRUE , 
       test.subgroup = FALSE , 
       fs.hetstat = 11,
       print.tau2 = TRUE ,
       backtransf = TRUE ,
       print.Q = TRUE , 
       print.pval.Q = TRUE , 
       print.I2 = TRUE , 
       rightcols = FALSE , 
       xlim = c(0 ,1),
       leftcols = c("studlab",  "effect",  "ci"), 
       leftlabs = c("Study",  "Proportion",  "95% C.I."), 
       text.random.w = "Subgroup proportion", 
       text.random = "Overall proportion", 
       xlab = "Antibiotics dispensed for URTI", 
       pscale = 1, 
       smlab = " ", 
       weight.study = "random", 
       squaresize = 0.5, 
       col.square = "black", 
       col.diamond = "black",
       col.diamond.lines = "black", 
       digits = 2,  
       spacing = 1.1)
dev.off()
```

